---
title: "Impact of lifestyle in chronic pain and depression: a Propensity Score Analysis in UK Biobank Extended Tables"
output: 
  word_document:
    reference_docx: ~/Desktop/PhD/projects/UKB_CP_MDD_brain_structure/writeup/RMD_template.docx
---

```{r setup, include=FALSE}
library(kableExtra)
library(dplyr)
library(flextable)
library(stringr)
library(tidyr)
library(knitr)
library(openxlsx)
```

# Table of contents:\

## 1. Supplementary Methods
### 1.1. Treatment, outcome and matching variables in UK Biobank

## 2. Supplementary Results
### 2.1. Observation Tables
### 2.2. Balance Tables
### 2.3. Data missingness
### 2.4. Results from main analysis

\newpage
# 1. Supplementary Methods

\n
## 1.1. Treatment, outcome and matching variables in UK Biobank

```{r Variable table, echo=F, results='asis'}
table_counter <- 1

variable_table <- read.xlsx(xlsxFile = "~/Desktop/PhD/projects/UKB_CP_MDD_lifestyle_counterfactual/writeup/variable_table.xlsx", fillMergedCells = TRUE, colNames = TRUE)

table <- flextable(variable_table) %>%
  merge_v(j = ~ Description + Field.ID + Coding) %>% ## Merge cells
  autofit() %>%
  fontsize(size = 10) %>%
  merge_h() %>%
  bg(i = c(1, 18, 37, 41), bg = "lightgrey", part = "body") %>%
  theme_booktabs() %>%
  vline() %>%
  hline()

## Function to fit table to page
FitFlextableToPage <- function(ft, pgwidth = 7){
  ft_out <- ft %>% autofit()
  ft_out <- width(ft_out, width = dim(ft_out)$widths*pgwidth /(flextable_dim(ft_out)$widths))
  return(ft_out)
}

FitFlextableToPage(table)

cat(paste0("Table S", table_counter, ". Table of matching, outcome and exposure variables used in UK Biobank"))
```


\newpage
# 2. Supplementary Results

\n
## 2.1. Observation Tables

```{r Observation tables, echo=F, results='asis'}

observation_table_files <- list.files(path = "~/Desktop/PhD/projects/UKB_CP_MDD_lifestyle_counterfactual/output/balancing/", pattern = "_observation_table.csv")

bal_table_files <- list.files(path = "~/Desktop/PhD/projects/UKB_CP_MDD_lifestyle_counterfactual/output/balancing/", pattern = "bal_table")

## Sort table lists so order is: full, female, male
observation_table_files_sorted <- c(grep("_full_", observation_table_files, value = T), grep("_female_", observation_table_files, value = T), grep("_male_", observation_table_files, value = T))

bal_table_files_sorted <- c(grep("_full_", bal_table_files, value = T), grep("_female_", bal_table_files, value = T), grep("_male_", bal_table_files, value = T))


## Observation tables
for (file in observation_table_files_sorted){
  
  table_counter <- table_counter + 1
  
  table <- read.csv(paste0("~/Desktop/PhD/projects/UKB_CP_MDD_lifestyle_counterfactual/output/balancing/", file ))
  names(table) <- c("Sample", "Non-exposure", "Exposure")
  
  f <- flextable(table) %>%
  autofit() %>%
  fit_to_width(7) %>%
  theme_booktabs()
   
  ## Get sample
   if (grepl("female", file)){
    sample <- "female sample"
  } else if (grepl("male", file)) {
    sample <- "male sample"
  } else if (grepl("full", file)) {
    sample <- "full sample"
  }
  
  ## Get exposure
  if (grepl("PA_low_", file)){
   exposure <- "physical activity" 
  } else if (grepl("insufficient_sleep_", file)){
    exposure <- "insufficient sleep" 
  } else if (grepl("high_alcohol_consumption_", file)){
    exposure <- "high alcohol consumption" 
  } else if (grepl("smoking", file)){
    exposure <- "smoking" 
  } else if (grepl("obese_", file)){
    exposure <- "obese" 
  } else if (grepl("lonely_", file)){
    exposure <- "loneliness" 
  } else if (grepl("unhealthy_diet_", file)){
    exposure <- "unhealthy diet" 
  }
  
  flextable_to_rmd(f)
  cat(paste0("Table S", table_counter, ". Observation table of ", exposure, " in ", sample,". Sample sizes averaged over multiple imputations."))
  cat("\n\n\n\n")
  
}

```

\n
## 2.2. Balance Tables
```{r Balance tables, echo=F, results='asis'}

## Balance tables
for (file in bal_table_files_sorted){
  
  table_counter <- table_counter + 1
  
  table <- read.csv(paste0("~/Desktop/PhD/projects/UKB_CP_MDD_lifestyle_counterfactual/output/balancing/", file ))
  names(table)[1] <- "Variables"
  
  ## Round numbers
  table <- table %>% mutate_if(is.numeric, round, digits=3)

  f <- flextable(table) %>%
  autofit() %>%
  fit_to_width(7) %>%
  theme_booktabs()
   
  ## Get sample
   if (grepl("female", file)){
    sample <- "female sample"
  } else if (grepl("male", file)) {
    sample <- "male sample"
  } else if (grepl("full", file)) {
    sample <- "full sample"
  }
  
  ## Get exposure
  if (grepl("PA_low_", file)){
   exposure <- "physical activity" 
  } else if (grepl("insufficient_sleep_", file)){
    exposure <- "insufficient sleep" 
  } else if (grepl("high_alcohol_consumption_", file)){
    exposure <- "high alcohol consumption" 
  } else if (grepl("smoking", file)){
    exposure <- "smoking" 
  } else if (grepl("obese_", file)){
    exposure <- "obese" 
  } else if (grepl("lonely_", file)){
    exposure <- "loneliness" 
  } else if (grepl("unhealthy_diet_", file)){
    exposure <- "unhealthy diet" 
  }
  
  flextable_to_rmd(f)
  cat(paste0("Table S", table_counter, ". Balance table of ", exposure, " in ", sample))
  cat("\n\n\n\n")
  
}

```

\newpage
## 2.3. Data Missingness

```{r data missingness, echo=F, results='asis'}
data_missingness <- read.csv("~/Desktop/PhD/projects/UKB_CP_MDD_lifestyle_counterfactual/output/data_missingness.csv")

data_missingness <- data_missingness %>%
  filter(X != "f.eid") %>%
   mutate(Variable = recode(X, ## Rename variables
                     followup_chronic_pain = "chronic pain - outcome",
                     followup_depression = "depression - outcome",
                     comorbid_CPDep = "comorbidity group - outcome",
                     PA_low = "low physcial activity - exposure",
                     insufficient_sleep = "insufficient sleep - exposure",
                     lonely = "loneliness - exposure",
                     smoking = "smoking - exposure",
                     high_alcohol_consumption = "high alcohol consumption - exposure",
                     obese = "obese - matching",
                     unhealthy_diet	= "unhealthy diet - matching",
                     PA_low_baseline =  "low physcial activity - matching",
                     insufficient_sleep_baseline = "insufficient sleep - matching",
                     lonely_baseline = "loneliness - matching varialble",
                     smoking_baseline = "smoking - matching",
                     high_alcohol_consumption_baseline =  "high alcohol consumption - matching",
                     obese_baseline = "obese - matching",
                     unhealthy_diet_baseline = "unhealthy diet - matching",
                     age = "age - matching",
                     sex = "sex - matching",
                     education = "education - matching",
                     employment = "employment - matching",
                     deprivation = "deprivation - matching",
                     general_health = "general health - matching",
                     living_with_partner = "living with partner - matching",
                     chronic_pain_baseline = "chronic pain - matching",
                     depression_baseline= "depression - matching")) %>%
  mutate(variable_type = case_when( ## Extract category
    grepl("outcome", Variable) ~ "Outcome",
    grepl("exposure", Variable) ~ "Exposure",
    grepl("matching", Variable) ~ "Matching"
  )) %>%
  arrange(variable_type, Variable) %>% ## Arrange variables
  rename("Missingness (%)" = "missingness....") %>%
  select(Variable, `Missingness (%)`)

f <- flextable(data_missingness) %>%
autofit() %>%
fit_to_width(7) %>%
theme_booktabs()

flextable_to_rmd(f)
table_counter <- table_counter + 1
cat(paste0("Table S", table_counter, ". Data missingness in full eligible sample"))

```


\newpage
## 2.4. Results from main analysis

```{r Results, echo=F, results='asis', warning=F}

results_table_files <- list.files(path = "~/Desktop/PhD/projects/UKB_CP_MDD_lifestyle_counterfactual/output/", pattern = "results.csv")

results_table_files_sorted <- c(grep("_full_", results_table_files, value = T), grep("_female_", results_table_files, value = T), grep("_male_", results_table_files, value = T))

single_disorder_table <- data_frame()
comorbidity_table <- data_frame()

## Observation tables
for (file in results_table_files_sorted){
  
  table <- read.csv(paste0("~/Desktop/PhD/projects/UKB_CP_MDD_lifestyle_counterfactual/output/", file ))

  ## Remove sensitivity analysis
  table <- subset(table, table$exposure != "sensitivity_PA_low")
  
  ## Add astriks to significant beta
  temp <- table$Coefficient.Estimate
  table$Coefficient.Estimate = ifelse(table$p_adjust < 0.05, paste0(temp, "*"), table$Coefficient.Estimate)
  table$Coefficient.Estimate = ifelse(table$p_adjust < 0.01, paste0(temp, "**"), table$Coefficient.Estimate)
  table$Coefficient.Estimate = ifelse(table$p_adjust < 0.001, paste0(temp, "***"), table$Coefficient.Estimate)

  ## Round number
  round_with_asterisks <- function(x, digits = 3) {
    num_part <- as.numeric(gsub("\\*", "", x))
    stars <- gsub("[0-9.-]", "", x)
    paste0(round(num_part, digits), stars)
  }
  table$Coefficient.Estimate <- round_with_asterisks(table$Coefficient.Estimate)

  ## Get sample
   if (grepl("female", file)){
    sample <- "Female Sample"
  } else if (grepl("male", file)) {
    sample <- "Male Sample"
  } else if (grepl("full", file)) {
    sample <- "Full Sample"
  }
  
  ## Combine all samples into single dataframe
   if (grepl("CP_Dep_", file)){
     
     ## Simplify and rename columns
       table <- table %>%
         select(Term, Coefficient.Estimate, outcome) %>%
         rename(Exposure = Term,
                {{sample}}:=Coefficient.Estimate)
     
     if(length(single_disorder_table) == 0){
       single_disorder_table <- table
       } else{
         single_disorder_table <- left_join(single_disorder_table, table, by = c("Exposure", "outcome"))
       }
   } else{
     
     ## Simplify and rename columns
     table <- table %>%
       select(Term, exposure, Coefficient.Estimate) %>%
       rename(`Nominal Group` = Term,
              Exposure = exposure,
              {{sample}}:=Coefficient.Estimate)

          if(length(comorbidity_table) == 0){
       comorbidity_table <- table
       } else{
         comorbidity_table <- left_join(comorbidity_table, table, by = c("Exposure", "Nominal Group"))
       }
   }
}
 

## Change variable names for display
single_disorder_table$Exposure[single_disorder_table$Exposure == "PA_low"] <- "physical activity"
single_disorder_table$Exposure[single_disorder_table$Exposure == "insufficient_sleep"] <- "insufficient sleep" 
single_disorder_table$Exposure[single_disorder_table$Exposure == "high_alcohol_consumption"] <- "high alcohol consumption" 
single_disorder_table$Exposure[single_disorder_table$Exposure == "smoking"] <- "smoking"
single_disorder_table$Exposure[single_disorder_table$Exposure == "obese"] <- "obese"
single_disorder_table$Exposure[single_disorder_table$Exposure == "loneliness"] <- "loneliness"
single_disorder_table$Exposure[single_disorder_table$Exposure == "unhealthy_diet"] <- "unhealthy diet"

comorbidity_table$Exposure[comorbidity_table$Exposure == "PA_low"] <- "physical activity"
comorbidity_table$Exposure[comorbidity_table$Exposure == "insufficient_sleep"] <- "insufficient sleep" 
comorbidity_table$Exposure[comorbidity_table$Exposure == "high_alcohol_consumption"] <- "high alcohol consumption" 
comorbidity_table$Exposure[comorbidity_table$Exposure == "smoking"] <- "smoking"
comorbidity_table$Exposure[comorbidity_table$Exposure == "obese"] <- "obese"
comorbidity_table$Exposure[comorbidity_table$Exposure == "loneliness"] <- "loneliness"
comorbidity_table$Exposure[comorbidity_table$Exposure == "unhealthy_diet"] <- "unhealthy diet"
  
## Split individual disorder results into seperate groups
CP_table <- single_disorder_table %>%
  filter(outcome == "followup_chronic_pain") %>%
  select(-outcome)

MDD_table <- single_disorder_table %>%
  filter(outcome == "followup_depression") %>%
  select(-outcome)
  
```

```{r CP Results, echo=F, results='asis', warning=F}
## Display imagees
flextable(CP_table) %>%
  autofit() %>%
  fit_to_width(7) %>%
  bold(i = ~ grepl("\\*", CP_table$`Full Sample`), j = "Full Sample") %>%
  bold(~ grepl("\\*", CP_table$`Female Sample`), j = "Female Sample") %>%
  bold(~ grepl("\\*", CP_table$`Male Sample`), j = "Male Sample")
  
table_counter <- table_counter + 1

```
Table S`r table_counter` Full results for association between lifestyle-related variables and chronic pain in full, female and male sample. \*, \*\* and \*\*\* represent significant results at p~FDR~\<.05, p~FDR~\<.01 and p~FDR~\<.001, respectively. Significant results are highlighted in bold.

\newpage

```{r Dep Results, echo=F, results='asis', warning=F}
## Display imagees
flextable(MDD_table) %>%
  autofit() %>%
  fit_to_width(7) %>%
  bold(i = ~ grepl("\\*", MDD_table$`Full Sample`), j = "Full Sample") %>%
  bold(~ grepl("\\*", MDD_table$`Female Sample`), j = "Female Sample") %>%
  bold(~ grepl("\\*", MDD_table$`Male Sample`), j = "Male Sample")
  
table_counter <- table_counter + 1
```
Table S`r table_counter` Full results for association between lifestyle-related variables and depression in full, female and male sample. \*, \*\* and \*\*\* represent significant results at p~FDR~\<.05, p~FDR~\<.01 and p~FDR~\<.001, respectively. Significant results are highlighted in bold.

\newpage

```{r Comorbidity Results, echo=F, results='asis'}
## Display imagees
flextable(comorbidity_table) %>%
  autofit() %>%
  fit_to_width(7) %>%
  bold(i = ~ grepl("\\*", comorbidity_table$`Full Sample`), j = "Full Sample") %>%
  bold(~ grepl("\\*", comorbidity_table$`Female Sample`), j = "Female Sample") %>%
  bold(~ grepl("\\*", comorbidity_table$`Male Sample`), j = "Male Sample")
  
table_counter <- table_counter + 1
```
Table S`r table_counter` Full results for association between lifestyle-related variables and nominal comorbidity groups in full, female and male sample. \*, \*\* and \*\*\* represent significant results at p~FDR~\<.05, p~FDR~\<.01 and p~FDR~\<.001, respectively. Significant results are highlighted in bold. 


